Gaia DR2 view of the GD-1 stellar stream

### 2018-04-28 paper/gd1.tex
Story
abstract edits

% typography
\setlength{\parindent}{1.\baselineskip}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\package}[1]{\textsl{#1}}
\newcommand{\gaia}{\textsl{Gaia}}
\newcommand{\DR}{\acronym{DR2}}

@@

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters leave behind thin, dynamically-cold streams
of stars that are extremely valuable tracers of the large- and small-scale
properties of mass around the Galaxy.
Many thin stellar streams have been discovered around the Milky Way, the
majority of which exist in the Galactic halo and are therefore primarily
sensitive to the distribution of dark matter.
Here we present astrometric data for one of the longest stellar streams --- the
``GD-1'' stream --- from data release 2 (\DR) of the \gaia\ mission.
We re-discover significant density variations along the stream by selecting
candidate member stars first using only the exquisite \gaia\ data.
Combined with filtering using Pan-STARRS photometry, we see clear evidence of
multiple prominent under-densities and haps in the stream.
These density variations --- and especially the gaps --- are possible
indications of past interactions with large, massive perturbers.
Given that the orbit of the stream crosses the Milky Way disk at large radius,
these perturbers are likely halo objects and have masses consistent with XX,
suggestive of dark matter sub-halos.
\end{abstract}

@@

Dynamically cold stellar streams, likely formed from the tidal disruption of
globular clusters, have been discovered throughout the Milky Way halo (e.g.,
\citealt{TODO}).

To date, $\approx$13--14 thin streams have been discovered by searching large-area sky surveys such as the Sloan Digital Sky Survey (SDSS) and Pan-STARRS1 survey (PS1). These previously known thin streams span a wide range of lengths \citep[$\sim$$2$--$60$ deg;][]{bernard14,grillmair06} and masses \citep[$\sim$$10^3$--$10^5~\msun$;][]{TODO,TODO}, however, only two such streams have clear progenitor systems: NGC 5466 and Palomar 5 \citep{TODO,TODO}. Though it has been shown with many independent methods that modeling these thin streams can place strong constraints on the gravitational potential in which they form \citep{apw14,TODO}, many efforts to apply these models to real streams around the Milky Way have been hampered by the small number of known progenitor systems.

@@

- Transform to GD-1 coordinates (Koposov et al.)
    - Can see deviation from great circle
- Cut to be close to latitude = 0
- Cut to get proper motion over-density
- Color-magnitude filtering

Figure 1:
- PM cut (panel a)
- PM + CMD cut (panel b)
- SFD (panel c)??

Comment on these here - add figure for both in Appendix?
- Dust / no significant extinction - combine with other figure
- Scanning pattern - just discuss, no figure

Most of the clearly identifiable portions of the GD-1 stream are located at high
Galactic latitudes ($b > 35^\circ$), and we therefore do not expect significant
dust extinction or variations in extinction along the stream.
\figurename~\ref{fig:sfd} shows the $V$-band extinction in the region around the
GD-1 stream, computed from the Schlegel-Finkbeiner-Davis extinction map
(\cite{Schlegel:1998}; hereafter SFD).
Vertical lines indicate the two regions clearly identifiable as under-densities
from the proper-motion-selected stream members.
The maximum $A_V$ is $\approx$0.07 mag in both the under-density and gap
regions.
The dispersion in $A_V$ in the region $-60^\circ < \phi_1 < 0^\circ$ and
$-1^\circ < \phi_1 < 1^\circ$ is $\std_{A_V} \approx 0.03~\textrm{mag}$.

% Notebook: GD1-dust-and-completeness
\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{sfd.pdf}
\end{center}
\caption{%
Colored background shows the \gaia\ $G$-band extinction, $A_G$, in the GD-1
coordinate system.
Vertical lines roughly show the regions identified as the under-density (UD) and
gap.
The maximum extinction in the UD or Gap regions is $\approx$0.07 mag.
\label{fig:sfd}
}
\end{figure}

@@

\acknowledgements{
thanks: Belokurov, Casey, Geha, Hogg, Johnston, Koposov, Schlafly, Spergel
This research was started at the NYC Gaia DR2 Workshop at the Center for Computational Astrophysics of the Flatiron Institute in 2018 April.
AB acknowledges generous support from the Institute for Theory and Computation at Harvard University.
All code used in this work and all results are available at \url{https://github.com/adrn/GD1-DR2}.
}

>>>>

% typography
\setlength{\parindent}{1.\baselineskip}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\package}[1]{\textsl{#1}}
\newcommand{\gaia}{\textsl{Gaia}}
\newcommand{\pans}{\textsl{PanSTARRS}}
\newcommand{\DR}{\acronym{DR2}}
\newcommand{\msun}{\rm M_\odot}

@@

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters are transformed into thin, dynamically-cold streams
of stars that are extremely valuable tracers of the large- and small-scale
properties of mass around the Galaxy.
Most of such streams discovered in the Milky Way reside in its halo, and are therefore primarily sensitive to the distribution of dark matter.
% Many thin stellar streams have been discovered around the Milky Way, the
% majority of which exist in the Galactic halo and are therefore primarily
- cleanest sample, foreground removed w gaia parallaxes; members selected on combination of gaia pm and panstarrs colors
- we thus got the crispest view of the stream to date, revealing in high resolution the rich structure which was only hinted at in the previous generation of data
% We present astrometric data for one of the longest stellar streams --- the
% GD-1 stream --- from the second data release (\DR) of the \gaia\ mission.
% We re-discover significant density variations along the stream by selecting
% candidate member stars using the exquisite \gaia\ data alone, including the
% existence of a prominent, apparent gap in the stream.
% <Using photometric data from the \pans, we refine our selection ...>
These density variations --- and especially the gap --- are possible indications
of past interactions with large, massive perturbers.
<estimate the mass of interaction...>
\end{abstract}

@@

Dynamically cold stellar streams, likely formed from the tidal disruption of globular clusters, are sensitive to the large and small scale distribution of matter in the halo

Thin stellar streams form from the tidal disruption of low-dispersion stellar systems such as globular clusters \citep[e.g.,][]{TODO}. To date, $\approx$13--14 thin streams have been discovered by searching large-area sky surveys such as the Sloan Digital Sky Survey (SDSS) and Pan-STARRS1 survey (PS1). These previously known thin streams span a wide range of lengths \citep[$\sim2$--$60$ deg;][]{bernard14,grillmair06} and masses \citep[$\sim10^3$--$10^5~\msun$;][]{TODO,TODO}, however, only two such streams have clear progenitor systems: NGC 5466 and Palomar 5 \citep{TODO,TODO}. Though it has been shown with many independent methods that modeling these thin streams can place strong constraints on the gravitational potential in which they form \citep{apw14,TODO}, many efforts to apply these models to real streams around the Milky Way have been hampered by the small number of known progenitor systems.

@@

Most of the clearly identifiable portions of the GD-1 stream are located at high
Galactic latitudes ($b > 35^\circ$), and we therefore do not expect significant
dust extinction or variations in extinction along the stream.
\figurename~\ref{fig:sfd} shows the $V$-band extinction in the region around the
GD-1 stream, computed from the Schlegel-Finkbeiner-Davis extinction map
(\cite{Schlegel:1998}; hereafter SFD).
Vertical lines indicate the two regions clearly identifiable as under-densities
from the proper-motion-selected stream members.
The maximum $A_V$ is $\approx$0.07 mag in both the under-density and gap
regions.
The dispersion in $A_V$ in the region $-60^\circ < \phi_1 < 0^\circ$ and
$-1^\circ < \phi_1 < 1^\circ$ is $\sigma_{A_V} \approx 0.03~\textrm{mag}$.

% Notebook: GD1-dust-and-completeness
\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{sfd.pdf}
\end{center}
\caption{%
Colored background shows the $V$-band extinction, $A_V$, in the GD-1
coordinate system.
Vertical lines roughly show the extents the identified under-density (UD) and
gap.
The maximum extinction in the UD or Gap regions is $\approx$0.07 mag.
\label{fig:sfd}
}
\end{figure}

@@

\acknowledgements{
Gaia
Belokurov, Casey, Geha, Hogg, Johnston, Koposov, Lisanti, Schlafly, Spergel
This research was started at the NYC Gaia DR2 Workshop at the Center for Computational Astrophysics of the Flatiron Institute in 2018 April.
AB acknowledges generous support from the Institute for Theory and Computation at Harvard University.
All code used in this work and all results are available at \url{https://github.com/adrn/GD1-DR2}.
}
